Gazing at the Robot's Head Reflects our Tendency to Anthropomorphise

### 2016-09-06 main.tex
WIP discussion

\newcommand{\anti}{{$\mathcal{A}_{start}$\xspace}}
\newcommand{\antf}{{$\mathcal{A}_{end}$\xspace}}
\newcommand{\deltaant}{{$\Delta_{\mathcal{A}}$\xspace}}

@@

    \item the mere observation of a robot acting in an \emph{human-like}
        situation suffice to increase the level of
        anthropomorphic attributions to robots \emph{in general} (\h{4}).
\end{itemize}

@@

This result also supports \h{4}: cognitive priming impacts the
attribution of anthropomorphic features to robots beyond the specific Nao robot
shown in the stimuli. It appears that participants transfer to some extent the
cognitive capabilities that they ascribe to the Nao robot in the
\emph{human-like} cognitive condition to robots in general. We can not, however,
conclude on any possible \emph{lasting} anthropomorphic ascriptions to robots as we
measure the anthropomorphic perception immediately after the experiment.

@@

We found a significant negative correlation between
$\delta_{\mathcal{H},\mathcal{R}}^{\text{gaze}}$ and
the initial tendency to anthropomorphise \anti{} (Pearson Correlation Coefficient $r(54) = -0.43,
p < .001$, figure~\ref{h2}). The higher the (initial) tendency to anthropomorphise, the smaller
the difference in gazing behaviours between the robot and human conditions. This
supports hypothesis \h{2}.

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{H2}\label{GazeDifference-vs-ICA}
    \caption{Gaze distribution difference $\delta_{\mathcal{H},
    \mathcal{R}}^{\text{gaze}}$ \vs initial tendency to anthropomorphise \anti{}}
    \label{h2}
\end{figure}

@@

Our experimental protocol enabled us to examine two additional questions: by
administrating the Godspeed questionnaire on anthropomorphism and mind
attribution first as pre-test, we gained a picture of the initial tendency of
our participants to anthropomorphise robots. Using our \deltaant measure of the
gazing behaviour similarity, we have evidenced that people with an higher
tendency to anthropomorphise robots, also exhibit a more similar gazing
behaviour when they watch either a human or a robot performing the same task. If
confirmed (in particular with balanced human genders), this is a significant
result. The Godspeed questionnaire assesses how likely we are to
anthropomorphise machines in an abstract way: no concrete robots nor interaction
situations are shown; the participants might react to the sentence ``Please rate
your impression of robots'' in widely different ways depending on their own
constructed images of what robots are. And yet, our results show a significant
correlation with gazing behaviour: the more you tend to anthropomorphise, the
more you look at robots the way you look at humans. This implies that the mere
propensity to attribute physical or mental human traits to robots suffice to
influence a mostly unconscious (yet easy to observe) physical behaviour.  This
link between psychological attributions and a resulting physical behaviour might
be of interest to the Human-Robot Interaction community, especially if we endow
the robot with the ability to measure itself, \emph{during} the interaction, the
time spent by the human on its face. This would for instance allow the robot to
build an initial model of the cognitive expectations held by the human regarding
the robot.

Hypothesis \h{2} looked at a possible interaction between our tendency to
anthropomorphise (as measured in the pre-questionnaire) and our actual gaze
behaviour, the hypothesis being that high anthropomorphisers would show smaller
differences between the way they look at robots and humans (since they would
consider robots closer to humans than low anthropomorphisers). Our data supports
this hypothesis: participants reporting an higher tendency to anthropomorphise
tend to have closer gazing behaviours when they look at humans or at robots:
statistically speaking, the tendency to anthropomorphise robots
is reflected in the way one looks at them: you look at them
like you would look at humans.

The third hypothesis \h{3} proposes that head fixations are related to the
cognitive expectations, which is clearly supported by our data: after human-like cognitive
priming, one tends to spend more time looking at the head of the interacting
agent than after machine-like cognitive priming. We also hypothesised that, in turn,
head fixations could act as a \emph{proxy to anthropomorphic projections}: as
underlined in the previous section, our data also support this.

Finally, the hypothesis \h{4} looks at the dynamic of anthropomorphic
projections, with the hypothesis that watching an interaction with a particular
robot after a human-like cognitive priming may impact \emph{globally} the
anthropomorphic perceptions of robots. This is supported by our study.

>>>>

\newcommand{\anti}{{$\mathcal{A}_{start}$\xspace}}
\newcommand{\antf}{{$\mathcal{A}_{end}$\xspace}}
\newcommand{\deltaant}{{$\Delta_{\mathcal{A}}$\xspace}}
\newcommand{\deltagaze}{{$\delta_{\mathcal{H},\mathcal{R}}^{\text{gaze}}$\xspace}}

@@

    \item the mere observation of a robot acting in an \emph{human-like}
        situation should not suffice to increase the level of
        anthropomorphic attributions to robots \emph{in general} (\h{4}).
\end{itemize}

@@

This result also contradicts \h{4}: cognitive priming impacts the
attribution of anthropomorphic features to robots beyond the specific Nao robot
shown in the stimuli. It appears that participants transfer to some extent the
cognitive capabilities that they ascribe to the Nao robot in the
\emph{human-like} cognitive condition to robots in general. We can not, however,
conclude on any possible \emph{lasting} anthropomorphic ascriptions to robots as we
measure the anthropomorphic perception immediately after the experiment.

@@

We found a significant negative correlation between
\deltagaze and
the initial tendency to anthropomorphise \anti{} (Pearson Correlation Coefficient $r(54) = -0.43,
p < .001$, figure~\ref{h2}). The higher the (initial) tendency to anthropomorphise, the smaller
the difference in gazing behaviours between the robot and human conditions. This
supports hypothesis \h{2}.

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{H2}\label{GazeDifference-vs-ICA}
    \caption{Gaze distribution difference \deltagaze \vs initial tendency to anthropomorphise \anti{}}
    \label{h2}
\end{figure}

@@

Our experimental protocol enabled us to examine two additional questions: by
administrating the Godspeed questionnaire on anthropomorphism and mind
attribution first as pre-test, we gained a picture of the initial tendency of
our participants to anthropomorphise robots. Using our \deltagaze measure of the
gazing behaviour similarity, we have evidenced that people with an higher
tendency to anthropomorphise robots, also exhibit a more similar gazing
behaviour when they watch either a human or a robot performing the same task. If
confirmed (in particular with balanced human genders), this is a significant
result. The Godspeed questionnaire assesses how likely we are to
anthropomorphise machines in an abstract way: no concrete robots nor interaction
situations are shown; the participants might react to the sentence ``Please rate
your impression of robots'' in widely different ways depending on their own
constructed images of what robots are. And yet, our results show a significant
correlation with gazing behaviour: the more you tend to anthropomorphise, the
more you look at robots the way you look at humans. This implies that the mere
propensity to attribute physical or mental human traits to robots suffice to
influence a mostly unconscious (yet easy to observe) physical behaviour.  This
link between psychological attributions and a resulting physical behaviour might
be of interest to the Human-Robot Interaction community, especially if we endow
the robot with the ability to measure itself, \emph{during} the interaction, the
time spent by the human on its face. This would for instance allow the robot to
build an initial model of the cognitive expectations held by the human regarding
the robot.

Lastly, we investigated one of the dynamics of anthropomorphic projections: is
watching a specific robot after a human-like cognitive priming sufficient to
impact the \emph{general} perception of robots as anthropomorphic agents. By
administrating the same Godspeed questionnaire as a post-test, we computed 
\deltaant to assess possible change in the participants' general tendency to
anthropomorphise. The stimuli were short videos of `toy' interaction scenarios,
and we did not anticipate any transfer of the anthropomorphic traits ascribed onto
the robot seen in the videos, to the broader tendency to anthropomorphise
robots in general, as assessed in the post-questionnaire. Our result challenge
this assumption: \deltaant is significantly higher for the participants who have been
primed with higher cognitive expectation before watching the robot performing
the tasks. This suggests a form of transfer: participants changed somewhat their
minds about robots being with however
